A fast and Low-Overhead Technique to Secure Programs Against Integer Overflows

### 2012-09-06 doc/Manuscript/CGO/paper.tex

% What are integer overflows.
The most popular programming languages, including C, C++ and Java, limit the
size of primitive numeric types.
For instance, the \texttt{int} type, in Java, ranges from $-2^{31}-1$ to
$2^{31}$.
Consequently, there exist numbers that cannot be represented by these types.
In general, these programming languages resort to a {\em wrapping-arithmetics}
semantics~\cite{Warren02} to perform integer operations.
If a number $n$ is too large to fit into a primitive data type $T$, then $n$'s
value wraps around, and $n \mbox{module} T_{max}$ wounds up represented
instead.
There are situations in which this semantics is acceptable~\cite{Dietz12}.
For instance, programers might rely on this behavior to implement hash functions
and random number generators.
On the other hand, there exist also situations in which this behavior might
lead a program to produce unexpected results.
As an example, in 1996, the Ariane 5 rocket was lost due to an arithmetic
overflow -- a bug that resulted in a loss of more than U\$370 million
dollars~\cite{Dowson97}.
% TODO We need to explain better the wrapping-arithmetics, and we need to cite it
% properly. What could be a good (official) citation?
% And we need to cite

@@

% Our experimental results.
We use our range analysis to reduce the runtime overhead imposed by a dynamic
instrumentation library.
This instrumentation framework, also a contribution of this paper, has been
implemented in the LLVM compiler~\cite{Lattner04}.
We have used it to log overflows in a vast number of programs, and in this
paper we focus on SPEC CPU 2006.
Our implementation has been able to reproduce the results obtained by
previous work.
We have re-discovered the integer overflows recently observed by
Dietz {\em et al.}~\cite{Dietz12}, and the performance of our instrumentation
library, without the support of range analysis, matches the state-of-the-art
implementation of Brumley {\em et al.}
Our range analysis algorithm has been able to remove XX\% of the overflow
checks created by the dynamic instrumentation framework.
All our experiments use an interprocedural version of range analysis.
We inline small functions to make our implementation more context sensitive.
Our experiments show that the speed and memory consumption of our range
analysis grows linearly with the program size.
By comparing the ranges that we determine statically with the results found by
a profiler, we see that more than 50\% of the bounds that we found are within a
factor of 50\% of the values that variables actually assume.

@@

\begin{tabular}{lcl}
$x + \infty = \infty, x \neq -\infty$ & \mbox{\hspace{0.1cm}} & $x - \infty = - \infty, x \neq +\infty$ \\
$x \times \infty = \infty$ if $x > 0$ & & $x \times \infty = -\infty$ if $x < 0$ \\
$0 \times \infty = 0$ & & $(-\infty) \times \infty = \ $ not defined $$ \\
\end{tabular}

@@

Given an interval $\iota = [l, u]$, we let $\lb{\iota} = l$, and
$\ub{\iota} = u$.
We let \varset{V} be a set of constraint variables, and
$I: \varset{V} \mapsto \varset{Z}^2$ a
mapping from these variables to intervals in $\varset{Z}^2$.
Our objective is to solve a constraint system $C$, formed by constraints such
as those seen in Figure~\ref{fig:eval_function}(left).
We let the $\phi$-functions be as defined by Cytron
{\em et al.}~\cite{Cytron91}: they join different variable names into a single
definition.
Figure~\ref{fig:eval_function}(right) defines a valuation function $e$ on the
interval domain.
Armed with these concepts, we define the range analysis problem as follows:

>>>>

% What are integer overflows.
The most popular programming languages, including C, C++ and Java, limit the
size of primitive numeric types.
For instance, the \texttt{int} type, in C++, ranges from $-2^{31}$ to
$2^{31}-1$.
Consequently, there exist numbers that cannot be represented by these types.
In general, these programming languages resort to a {\em wrapping-arithmetics}
semantics~\cite{Warren02} to perform integer operations.
If a number $n$ is too large to fit into a primitive data type $T$, then $n$'s
value wraps around, and $n \mbox{module} T_{max}$ wounds up represented
instead.
There are situations in which this semantics is acceptable~\cite{Dietz12}.
For instance, programers might rely on this behavior to implement hash functions
and random number generators.
On the other hand, there exist also situations in which this behavior might
lead a program to produce unexpected results.
As an example, in 1996, the Ariane 5 rocket was lost due to an arithmetic
overflow -- a bug that resulted in a loss of more than US\$370 million~\cite{Dowson97}.

% TODO We need to explain better the wrapping-arithmetics, and we need to cite it
% properly. What could be a good (official) citation?

@@

% Our experimental results.
We use our range analysis to reduce the runtime overhead imposed by a dynamic
instrumentation library.
This instrumentation framework, also a contribution of this paper, has been
implemented in the LLVM compiler~\cite{Lattner04}.
We have used it to log overflows in a vast number of programs, and in this
paper we focus on SPEC CPU 2006.
Our implementation has been able to reproduce the results obtained by
previous work.
We have re-discovered the integer overflows recently observed by
Dietz {\em et al.}~\cite{Dietz12}, and the performance of our instrumentation
library, without the support of range analysis, matches the state-of-the-art
implementation of Brumley {\em et al.}
Our range analysis algorithm has been able to remove 24.93\% of the overflow
checks created by the dynamic instrumentation framework.
All our experiments use an interprocedural version of range analysis.
We inline small functions to make our implementation more context sensitive.
Our experiments show that the speed and memory consumption of our range
analysis grows linearly with the program size.
By comparing the ranges that we determine statically with the results found by
a profiler, we see that more than 50\% of the bounds that we found are within a
factor of 50\% of the values that variables actually assume.

@@

\begin{tabular}{lcl}
$x + \infty = \infty$ & \mbox{\hspace{0.1cm}} & $x - \infty = - \infty, x \neq +\infty$ \\
$x \times \infty = \infty$ if $x > 0$ & & $x \times \infty = -\infty$ if $x < 0$ \\
$0 \times \infty = 0$ & & $(-\infty) \times \infty = \ $ not defined $$ \\
\end{tabular}

@@

Given an interval $\iota = [l, u]$, we let $\lb{\iota} = l$, and
$\ub{\iota} = u$, where $\lb{\iota}$ is the lower bound and $\ub{\iota}$ is
the upper bound of a variable.
We let \varset{V} be a set of constraint variables, and
$I: \varset{V} \mapsto \varset{Z}^2$ a
mapping from these variables to intervals in $\varset{Z}^2$.
Our objective is to solve a constraint system $C$, formed by constraints such
as those seen in Figure~\ref{fig:eval_function}(left).
We let the $\phi$-functions be as defined by Cytron
{\em et al.}~\cite{Cytron91}: they join different variable names into a single
definition.
Figure~\ref{fig:eval_function}(right) defines a valuation function $e$ on the
interval domain.
Armed with these concepts, we define the range analysis problem as follows:
